### 2012-10-11 3.tex
Completion of draft of chapter 3

\citet{Chomsky1965} argue that speakers internalize language-specific generalizations about possible and impossible words, but it is certainly possible to imagine otherwise. \citet[][320]{Zimmer1969} writes that phonotactic generalizations ``have no observable consequences in the course of the normal use of the language'', and \citeauthor{PE} echo this sentiment more recently:

\begin{quote}
Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for. \citep[][18]{PE}
\end{quote}

This represents a principled null hypothesis, but can be quickly dismissed in light of speaker's ready judgements of possible and impossible words observed by \citeauthor{Chomsky1965}. As shown in Chapter \ref{gradience}, wordlikeness tasks implicate prosodic wellformedness and cannot be reduced to purely statistical generalizations. Metalinguistic judgements are not the only task in which phonotactic knowledge is on display; speakers are also thought to make use \emph{possible word constraint} in more quotidien tasks like recognizing words in running speech \citep[e.g.,][]{Brown1956,McQueen1998b,Norris1997}.

struct/act
copenhagen
pure statistical critique

\citet{Fischer-Jorgensen1952} and \citet{Vogt1954}
note the inherent difficulty of distinguishing between structural and accidental phonotactic gaps.

Even if the criteria differentiating structural and accidental generalizations were purely statistical, an accidental generalization is one which lacks an antecedent cause, and it falls to the linguist to delimit a theory of these causes.

\citeauthor{Chomsky1965} were not the first linguists to take an interest in possible wordhood.

The early generativists identify these causes in mentalism, grounding the study of accidental gaps in what patterns are internalized, and ultimately how they are learned and revealed in psycholinguistic tasks. Consequently, for the generativist an accidental phonotactic generalization can only be one which speakers do not learn, do not attend to, and thus a generalization which is without mental antecedent.

Much of the latter-day literature on phonotactic knowledge, while adopting the representational machinery of generative phonology, expresses limited interest in what generalizations speakers internalize. In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique, a chi-square test, in an attempt to validate a generalization about Arabic root consonant co-occurrence facts, observations originally due to \citet{Greenberg1950}. The assumption appears to be that statistically valid static co-occurence restrictions will be reflected by wordlikeness judgements. While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies in the \citeauthor{McCarthy1988}ian model; rather, the phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Dmitrieva2008a,Dmitrieva2008b,Elmedlaoui1995,Graff2011,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}. In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} makes the tacit assumption of this approach explicit, claiming that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers.

\begin{quote}
\ldots{} the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
\end{quote}

There are several arguments against this inference. First, linguists since Saussure have noted diachronic causes for statistical generalizations which lack any synchronic reality. For instance, Saussure \citeyear[202f.]{CLG} recognizes the a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}, and that this makes intervocalic \emph{s} rare in Classical Latin. However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{Gorman2012e} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.} Secondly, wordlikeness generalizations are constrained by negative results in learnability (e.g., \citealp{Gold1967}; see \citealt{Yang2012} for a recent review), and may also be restricted by markedness constraints \citep[e.g.,][]{Becker2011,Hayes2009,Moreton2002}. Finally, there are a number of results which suggest patterns obvious to the linguist go unlearned \citep{Becker2011,Hayes2006,Zimmer1969}. This argues against a direct link between statistical criteria and the synchronic reality of phonotactic generalizations. Consequently, grammatical models which closely mimic lexical statistics but which lack psycholinguistic support (e.g., Gitksan: \citealt{Brown2010}; Muna: \citealt{Anttila2008a}, \citealt{Coetzee2008a}; Shona: \citealt[][385]{Hayes2008a}) may not be appropriate models of speakers' phonotactic knowledge.

This chapter is a comprehensive investigation of three well-known phonotactic generalizations in Turkish, as well as the results of a wordlikeness task designed to investigate the synchronic reality of these generalizations \citep{Zimmer1969}. This study merits reconsideration because there is widespread disagreement about the analysis of \citeauthor{Zimmer1969}, and prior discussions fail to relate the behavioral data to lexical statistics and competing formalizations of the constraints involved. One of these constraints, \textsc{Labial Attraction}, is shown to be highly statistically reliable but to be inactive in the wordlikeness task. This undermines the widespread use of statistical criteria to determine which of the vast number of phonotactic generalizations are internalized by speakers and which are not. The simplest hypothesis is that \textsc{Labial Attraction} is inactive in wordlikeness tasks because it is phonologically inactive a well.

@@

\citet{Lees1966b,Lees1966a} proposes three constraints on Turkish vowel sequences, constraints which are the focus of many subsequent studies. In this setion, these constraints are formalized, and where possible, related to phonological alternations and to behavioral evidence suggesting that speakers internalize restrictions. The following feature specification for the eight vowels of Turkish vowels is assumed throughout.

@@

Two non-conventional notations are used in this chapter. First, directional application conditions are used when necessary to model the directionality of harmony, rather than ad hoc contextual restrictions (e.g., that the trigger be aligned with the left of a prosodic word). Directional application provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}) and simplifies the analysis of harmony in Turkish \citep[209f.]{Anderson1974}. Further, \citet{Johnson1972} and \citet{Kaplan1994} prove that rules specified for directional application are formally learnable. Secondly, rather than the use of an unbounded number of Greek-letter variables ($\alpha$, $\beta$, etc.) over feature values $\{+, -\}$, only a single variable, denoted by `$=$', is used. Following \citet{McCawley1973}, a structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].\footnote{This is more restrictive than the Greek-letter variable notation, in that the value of $=$ for some feature F cannot be applied to F$'$ if F $\ne$ F$'$; \citet{Odden2012} dismisses the two potential counterexamples against this restrict, both adduced in \emph{SPE} (352-353).}

@@

If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllablic roots to contain only [$+$\textsc{Back}] or [$-$\textsc{Back}] vowels, a tendency which will be quantified below. \textsc{Backness Harmony} also triggers alternations in inflectional suffix vowels. For instance, the nominative plural (nom.pl.) suffix is \emph{-ler}, when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}].

\begin{example}[The Turkish nominative] \label{turknom}
\begin{tabular}{l l l l@{ }l}
   & \emph{nom.sg.} & \emph{nom.pl.} \\
a. & {ip}           & {ipler}    & `rope'         & \citep[][216]{Clements1982} \\
%   & {el}           & {eller}    & `hand'         \\
   & {köy}          & {köyler}   & `village'      \\
   & {yüz}          & {yüzler}   & `face'         \\
   & {kız}          & {kızlar}   & `girl'         \\
%   & {sap}          & {saplar}   & `stalk'        \\
%   & {son}          & {sonlar}   & `end'          \\
   & {pul}          & {pullar}   & `stamp'        \\
b. & {neden}        & {nedenler} & `reason'       & \citep{TELL} \\
   & {kiler}        & {kilerler} & `pantry'       \\ % front
   & {pelür}        & {pelürler} & `tissue paper' \\ % back
   & {boğaz}        & {boğazlar} & `throat'       \\ % back
   & {sapık}        & {sapıklar} & `pervert'      \\ % back
\end{tabular}
\end{example}

@@

\begin{example}[Exceptional Turkish nominatives] \label{turkexcept}
\begin{tabular}{l l l l@{ }l}
   & \emph{nom.sg.} & \emph{nom.pl.} \\
a. & {mezar}        & {mezarlar} & `grave' & \citep{TELL} \\
   & {model}        & {modeller} & `model' \\
   & {silah}        & {silahlar} & `weapon'     \\
   & {memur}        & {memurlar} & `bureaucrat' \\
   & {sabun}        & {sabunlar} & `soap'       \\
b. & {saat}         & {saatler}  & `hour, clock' & \citep{Goksel2005} \\
   & {harf}         & {harfler}  & `(alphabetic) letter' \\ 
   & {etol}         & {etoller}  & `fur stole' \\
\end{tabular}
\end{example}

\noindent
There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler} despite the fact that their final root vowel is [$+$\textsc{Back}]. While it is uncontroversial that the disharmonic suffixes of (\ref{turkexcept}b) are no more than very sporadic exceptions to \textsc{Backness Harmony}, root disharmony (\ref{turkexcept}a) has been the subject of much debate. As disharmonic roots still trigger suffix harmony, \citet[212, 289]{Anderson1974} and \citet{Iverson1978} distinguish suffix harmony alternations from the sequence structure constraint governing root harmony.

The disadvantage of this account is that it introduces ``duplication'' (in the sense of \citealt{Kisseberth1970b} and \citealt{Kenstowicz1977}) of sequence structure and phonological generalizations, differning only in their patterns of exceptionality. However, \citet[][197f.]{Zonneveld1978} shows that the theory of exceptionality proposed in \emph{SPE} can account for suffix harmony in disharmonic roots.\footnote{An isomorphic analysis is available within less restrictive theories of phonological exceptionality \citep[e.g.,][]{Kisseberth1970,Pater2009}.} \citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention. A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}]. In other words, no form is never truly an ``exception''; rather, some forms fail to match the extended structural description of $R$ which requires that the target be [$+R$]. If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement. Thus \textsc{Backness Harmony} is able to account for both root and suffix harmony.

It is necessary to dispense with an alternative analysis proposed by \citet{Clements1982} and \citet{Inkelas1997}. Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.  \emph{kul} `servant', \emph{kepek} `bran' vs. \emph{kapak} `lid'), whereas harmonic roots are those in which the backness of any remaining vowels is predictable. \citeauthor{Clements1982} propose that these vowels, as well as harmonizing suffix vowels, are underspecified for backness, whereas the non-initial vowels of disharmonic roots are fully specified. This is schematized below.

@@

One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being a Structure Preservation condition \citep{Kiparsky1985}. However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a perceptual default. A precedent for this is the surface-oriented interpretation of the tonal Obligatory Contour Principle proposed by \citet[134]{Goldsmith1976} and \citet{Odden1986}, under which adjacent identical tones are automatically attributed to a single underlying tone. However, this interpretation makes the underspecification analysis a mere notational variant of the rule exceptionality account, in which [$+$\textsc{Backness Harmony}] is the default.} On these grounds, the underspecification analysis is rejected here in favor of the rule exceptionality account.

While harmony in non-derived environments can be inferred from the aforementioned suffix alternations, no evidence has yet been presented to show internalize the tendency for roots to conform to backness harmony. If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it. Several other ``external'' facts demonstrate that this is not the case.

@@

\begin{example}[Variable non-native cluster adaptation (\citealp{Clements1982}:247)] 
\begin{tabular}{l l l l l l}
a. & {spiker}  & \alt{} & {sipiker}  & `announcer' \\
   & {fren}    & \alt{} & {firen}    & `break'     \\
b. & {trablus} & \alt{} & {tırablus} & `Tripoli'   \\
   & {kral}    & \alt{} & {kıral}    & `king'      \\
c. & {brom}    & \alt{} & {burom}    & `bromide'   \\
   & {prusya}  & \alt{} & {purusya}  & `Prussia'   \\
%b. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
%   & kredi   & \alt{} & kıredi   & `credit'    \\
\end{tabular}
\end{example}

@@

Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} This game is not indigenous to Turkish, but it is found in the grammar of the Turkic language Tuvan, where it is used to convey a sense of ``vagueness or jocularity''; \citeauthor{Harrison2001} report that it can be quickly taught to Turkish speakers. The game consists of reduplication of the base, with the first vowel of the reduplicant replaced with the [$+$\textsc{Back}] vowels \emph{a} or \emph{u}. Both in Tuvan and in the Turkish game, the second vowel of the reduplicant (shown in braces below) may also be effected by this process; when the base vowels are both [$-$\textsc{Back}], the insertion of a [$+$\textsc{Back}] vowel results in what \citeauthor{Harrison2001} call ``reharmonization'', shown in  (\ref{redupgame}a). On the other hand, as shown in (\ref{redupgame}b), disharmonic roots do not reharmonize.\footnote{A similar pattern is found in Tuvan, and in an unrelated Finnish language game \citep{Campbell1986}.}

\begin{example}[Turkish reduplication game (\citealp{Harrison2001}:231)] \label{redupgame}
\begin{tabular}{l l l l}
a. & {kibrit} & {kibrit}-\{{kabrıt}\} & `match'    \\
   & {bütün}  & {bütün}-\{{batın}\}   & `whole'    \\
b. & {mali}   & {mali}-\{{muli}\}     & `Mali'     \\
   & {butik}  & {butik}-\{{batik}\}   & `boutique' \\
\end{tabular}
\end{example}

@@

A number of studies have investigated the role of harmony in word-spotting tasks, though to reveal how speakers identify word boundaries in running speech. Many of these studies have been carried out in Finnish, which has a vowel harmony system similar to Turkish. \citet{Suomi1997} and \citet{Vroomen1998} task speakers with identifying harmonic disyllables in an auditory stream. Finnish speakers are better able to complete this task when the preceding syllable has a different backness specification than the disyllabic target. \citet{Kabak2010} report that Turkish \textsc{Backness Harmony} has the same effect on word-spotting tasks as it does in Finnish: Turkish speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by a disharmonic juncture (e.g., \emph{gölü-PAVO}) than when preceded by a harmonic juncture (e.g., \emph{golu-PAVO}), but this effect does not obtain for speakers of French, a language which lacks vowel harmony. These results imply that speakers of Turkish have internalized the tendency of harmonic sequences to be root-internal and of disharmonic transitions to mark word juncture.

The word-spotting experiment has been adapted for infants by \citep{Kampen2008}. In this study, 9-month-old infants are familiarized with an auditory stream containing harmonic disyllabic nonce words. At test time, the infants listen to the disyllabic nonce words in isolation using the head turn preference paradigm. Infants acquiring Turkish listen longer to nonce words preceded by a disharmonic juncture during familiarization (e.g., \emph{lo-NETIS}), whereas infants acquiring German, a language which lacks vowel harmony, do not exhibit this preference. Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic pseudowords such as \emph{paroz} over disharmonic pseudowords like \emph{nelok}, but German 6-month-old infants show no such preference.

@@

\subsubsection{Phonological description}

@@

\begin{example}[\textsc{Roundness Harmony} (condition: rightward application)]
$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~/~\begin{bmatrix}~=\textsc{Back}~\end{bmatrix}~\textrm{C}_0~\gap$
\end{example}

@@

\begin{example}[\textsc{Labial Attraction}]
$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\begin{bmatrix} -\textsc{Cons} \\ -\textsc{Rnd} \\ +\textsc{Back} \\ -\textsc{High} \end{bmatrix}~\textrm{C}_0~\begin{bmatrix} +\textsc{Cons} \\ +\textsc{Labial} \end{bmatrix}~\textrm{C}_0~\gap{}$
\end{example}

>>>>

There is a long tradition of distinguishing between structural phonotactic gaps and those which are simply accidental \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}. In purely statistical terms, an accidental gap is one which is likely to arise without any antecedent cause, i.e., by chance. The contribution of early generative discussions of this contrast \citep[e.g.,][]{Halle1962,Chomsky1965,SPE} is to ground the definitions of \emph{structural} and \emph{accidental} gaps in mentalism: an accidental gap is one which which lacks a mental antecedent because it goes unlearned by speakers.

Much of the latter-day literature on phonotactic knowledge, while adopting the representational machinery of generative phonology, takes less interest in the question of what generalizations speakers internalize. In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique to validate a generalization about Arabic root consonant co-occurrence facts originally due to \citet{Greenberg1950}. The assumption appears to be that statistically valid static co-occurrence restrictions will be reflected by wordlikeness judgements. While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies inspired by \citeauthor{McCarthy1988}'s study; in most cases, phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Elmedlaoui1995,Graff2011,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}. In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers:

\begin{quote}
\ldots{} the patterns outlined above are statistically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
\end{quote}

There are several arguments against this inference. First, linguists have long noted statistical generalizations which have a diachronic cause but which lack any synchronic reality. One of the earliest arguments of this type was put forth by Saussure (\citeyear[202f.]{CLG}), who recognizes the a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}, and that Classical Latin intervocalic \emph{s} is rare as a result. However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{Gorman2012e} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.} Similarly, [ʃ] in Modern English derives from Old English [sk] (e.g., \emph{fisc} 'fish'), and since long vowels before complex codas are not found in Old English, [ʃ] is rarely preceded by long vowels in word-final syllables to this day. \citet{Iverson2005} label this constraint ``phonologically accidental'', since a millennium of speakers have produced coinages (e.g., \emph{posh}, \emph{swoosh}) and borrowings (e.g., \emph{douche}, \emph{capiche}) in utter disregard of this generalization. Secondly, wordlikeness generalizations are absolutely constrained by negative results in learnability (see \citealt{Yang2012} for a recent review). Finally, there are a number of results which suggest patterns which were obvious to the linguist go unlearned \citep[e.g.,][]{Becker2011,Hayes2006,Hayes2009,HayesInPress}. This argues against a direct link between statistical criteria and the synchronic reality of phonotactic generalizations. The danger is that grammatical models which closely mimic lexical statistics but which lack psycholinguistic support (e.g., Gitksan: \citealt{Brown2010}; Muna: \citealt{Anttila2008a}, \citealt{Coetzee2008a}; Shona: \citealt[][385]{Hayes2008a}) may not match speakers' knowledge.

This chapter is a comprehensive investigation of three well-known phonotactic generalizations in Turkish using lexical statistics and the results of a wordlikeness task performed by \citet{Zimmer1969}. \citeauthor{Zimmer1969}'s experiments merit reconsideration because there is widespread disagreement concerning the results and prior discussions fail to relate the behavioral data to lexical statistics and to competing formalizations of the generalizations involved.

One of these generalizations, \textsc{Labial Attraction}, is shown to be highly statistically reliable but to be inactive in the wordlikeness task. This further undermines the widespread use of statistical criteria to determine which of the vast number of statistically reliable phonotactic generalizations are internalized by speakers and which are not. The simplest hypothesis is that \textsc{Labial Attraction} is inactive in wordlikeness tasks because it is phonologically inactive, corresponding to no phonological process in the language.

@@

\citet{Lees1966b,Lees1966a} proposes three constraints on Turkish vowel sequences, constraints which are the focus of many subsequent studies. In this section, these constraints are formalized, and where possible, related to phonological alternations and to behavioral evidence bearing on speakers' knowlege of the restrictions. The following feature specification for the eight vowels of Turkish vowels is assumed throughout.

@@

Two less familiar notations are used in this chapter. First, directional application conditions are used to derive the left-to-right, iterative properties of. Directional application explains provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt[209f.]{Anderson1974}, \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}). In fact, harmony in Turkish is one such case, insofar as simultaneous application makes the wrong predictions. Further, \citet{Johnson1972} and \citet{Kaplan1994} prove that rules specified for directional application are no more complex in the terms of formal learnability than that do not. Secondly, rather than the use of an unbounded number of Greek-letter variables ($\alpha$, $\beta$, etc.) over feature values $\{+, -\}$, only a single variable, denoted by `$=$', is used. Following \citet{McCawley1973}, a structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].\footnote{This is more restrictive than the Greek-letter variable notation, in that the value of $=$ for some feature F cannot be applied to F$'$ if F $\ne$ F$'$; \citet{Odden2012} dismisses well-known counterexamples against this restriction adduced in \emph{SPE} (352-353).}

@@

If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllabic roots to contain only [$+$\textsc{Back}] or [$-$\textsc{Back}] vowels, a tendency which will be quantified below. \textsc{Backness Harmony} also triggers alternations in inflectional suffix vowels. For instance, the nominative plural (nom.pl.) suffix is \emph{-ler}, when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}].

\begin{example}[The Turkish nominative]
\label{turknom}
\begin{tabular}{lllll}
   & \emph{nom.sg.} & \emph{nom.pl.} \\
a. & {ip}           & {ipler}    & `rope'         & \citep[][216]{Clements1982} \\
%  & {el}           & {eller}    & `hand'         \\
   & {köy}          & {köyler}   & `village'      \\
   & {yüz}          & {yüzler}   & `face'         \\
   & {kız}          & {kızlar}   & `girl'         \\
%  & {sap}          & {saplar}   & `stalk'        \\
%  & {son}          & {sonlar}   & `end'          \\
   & {pul}          & {pullar}   & `stamp'        \\
b. & {neden}        & {nedenler} & `reason'       & \citep{TELL} \\
   & {kiler}        & {kilerler} & `pantry'       \\
   & {pelür}        & {pelürler} & `tissue paper' \\
   & {boğaz}        & {boğazlar} & `throat'       \\
   & {sapık}        & {sapıklar} & `pervert'      \\
\end{tabular}
\end{example}

@@

\begin{example}[Exceptional Turkish nominatives] 
\label{turkexcept}
\begin{tabular}{lllll}
   & \emph{nom.sg.} & \emph{nom.pl.}             &                    \\
a. & {mezar}        & {mezarlar} & `grave'       & \citep{TELL}       \\
   & {model}        & {modeller} & `model'                            \\
   & {silah}        & {silahlar} & `weapon'                           \\
   & {memur}        & {memurlar} & `bureaucrat'                       \\
   & {sabun}        & {sabunlar} & `soap'                             \\
b. & {etol}         & {etoller}  & `fur stole'   & \citep{Goksel2005} \\
   & {saat}         & {saatler}  & `hour, clock' 	                  \\
%  & {harf}         & {harfler}  & `(alphabetic) letter' \\ 
\end{tabular}
\end{example}

\noindent
There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler} despite the fact that their final root vowel is [$+$\textsc{Back}]. While suffix disharmony appears to very sporadic, the co-occurrence of suffix harmony and root disharmony has been the subject of much debate. As disharmonic roots still trigger suffix harmony, 
%\citet[212, 289]{Anderson1974} and
\citet{Iverson1978} distinguish suffix harmony alternations from the sequence structure constraint governing root harmony.\footnote{\citet[29f.]{Kiparsky1968} discusses a parallel case in Finnish, and proposes a similar separation between the exception-filled sequence structure constraint and an exceptionless suffix harmony process. \citet[171f.]{Howard1972} correctly observes that this duplication is unnecessary.}

The disadvantage of this account is that it introduces ``duplication'' (in the sense of \citealt{Kisseberth1970b}) of sequence structure and phonological generalizations, differing only in their pattern of exceptionality. \citet[][197f.]{Zonneveld1978} shows that the theory of exceptionality proposed in \emph{SPE} can in fact account for suffix harmony in disharmonic roots.\footnote{An isomorphic analysis is available within less restrictive theories of phonological exceptionality \citep[e.g.,][]{Kisseberth1970,Pater2009}.} \citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention. A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}]. In other words, no form is never truly an ``exception''; rather, some forms fail to match the extended structural description of $R$ which requires that the target be [$+R$]. If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement. Thus \textsc{Backness Harmony} is able to account for both root and suffix harmony, and duplication is unnecessary.

It is necessary to dispense with an alternative proposed by \citet{Clements1982} and \citet{Inkelas1997} which reintroduces this duplication. Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.  \emph{kul} `servant', \emph{kepek} `bran' vs. \emph{kapak} `lid'), whereas harmonic roots are those in which the backness of any remaining vowels is predictable. \citeauthor{Clements1982} propose that these vowels, as well as harmonizing suffix vowels, are underspecified for backness, whereas the non-initial vowels of disharmonic roots are fully specified. This is schematized below.

@@

One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being a Structure Preservation condition \citep{Kiparsky1985}. However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a sort of default. A precedent for this is the surface-oriented interpretation of the tonal Obligatory Contour Principle proposed by \citet[134]{Goldsmith1976} and \citet{Odden1986}, under which adjacent identical tones are automatically attributed to a single underlying tone. However, this is merely a notational variant of the rule exceptionality account in which [$+$\textsc{Backness Harmony}] is the default.} On these grounds, the underspecification analysis is rejected here in favor of the rule exceptionality account.

While harmony in non-derived environments can be inferred from the aforementioned suffix alternations, no evidence has yet been presented to show internalize the tendency for roots to conform to backness harmony. If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it. Several other ``external'' facts suggest  that this is not the case, with the caveat that uncritical acceptance of hypotheses linking loanword adaptation and language games to the grammar is no way being implied.

@@

\begin{example}[Variable non-native cluster adaptation (\citealp{Clements1982}:247)] 
\begin{tabular}{lllll}
a. & {spiker}  & \alt{} & {sipiker}  & `announcer' \\
   & {fren}    & \alt{} & {firen}    & `break'     \\
b. & {trablus} & \alt{} & {tırablus} & `Tripoli'   \\
   & {kral}    & \alt{} & {kıral}    & `king'      \\
c. & {brom}    & \alt{} & {burom}    & `bromide'   \\
   & {prusya}  & \alt{} & {purusya}  & `Prussia'   \\
%b. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
%   & kredi   & \alt{} & kıredi   & `credit'    \\
\end{tabular}
\end{example}

@@

Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} 
The game is native to the related language Tuvan, where it is used to convey a sense of ``vagueness or jocularity''; it is not indigenous to Turkish, but can be taught quickly to children or adults. In this game, the base is reduplicated and the first vowel of the reduplicant replaced with a [$+$\textsc{Back}] vowel. In (\ref{redupgame}a), the second [$-$\textsc{Back}] vowel of the base is, in the reduplicant, ``reharmonized'' with the the inserted [$+$\textsc{Back}] vowel. The disharmonic roots of (\ref{redupgame}b) do not reharmonize.\footnote{A similar contrast between harmonic and disharmonic roots is found in Tuvan \citep{Harrison2001} and in an unrelated Finnish language game \citep{Campbell1986}.}

\begin{example}[Turkish reduplication game (\citealp{Harrison2001}:231)] 
\label{redupgame}
\begin{tabular}{llll}
a. & {kibrit} & {kibrit}-\{{kabrıt}\} & `match'    \\
   & {bütün}  & {bütün}-\{{batın}\}   & `whole'    \\
b. & {mali}   & {mali}-\{{muli}\}     & `Mali'     \\
   & {butik}  & {butik}-\{{batik}\}   & `boutique' \\
\end{tabular}
\end{example}

@@

A number of studies have investigated the role of harmony in word-spotting tasks, though to mimic auditory word recognition and segmentation in natural settings. Many of these studies have been carried out in Finnish, which has a vowel harmony system similar to that of Turkish. \citet{Suomi1997} and \citet{Vroomen1998} task Finnish speakers with identifying harmonic disyllables in an auditory stream. When the syllable preceding the disyllabic target has a different backness specification than the target, this faciliates speakers' recognition abilities. Presumably, disharmony facilitates the recognition of word boundaries. \citet{Kabak2010} find that Turkish \textsc{Backness Harmony} has a a similar effect: Turkish speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by a disharmonic juncture (e.g., \emph{gölü-PAVO}) than when preceded by a harmonic juncture (e.g., \emph{golu-PAVO}). \citet{Kabak2010} find that this effect does not obtain for speakers of French, a language which lacks vowel harmony. This implies that Turkish speakers have internalized the tendency of harmonic sequences to be root-internal and of disharmonic transitions to cross word boundaries.

The Turkish word-spotting experiment has been adapted for infants by \citet{Kampen2008}. In this study, 9-month-old infants are familiarized with an auditory stream containing harmonic disyllabic nonce words. At test time, the infants listen to the disyllabic nonce words in isolation using the head turn preference paradigm. Infants acquiring Turkish listen longer to nonce words preceded by a disharmonic juncture during familiarization (e.g., \emph{lo-NETIS}), whereas infants acquiring German, a language which lacks vowel harmony, do not exhibit this preference. Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic nonce words such as \emph{paroz} over disharmonic nonce words like \emph{nelok}, but German 6-month-old infants show no such preference.

@@

@@

\begin{example}[\textsc{Roundness Harmony} (condition: rightward application)]
$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Rnd} \end{bmatrix}~/~\begin{bmatrix}~=\textsc{Rnd}~\end{bmatrix}~\textrm{C}_0~\gap$
\end{example}

@@

\begin{example}[\textsc{Labial Attraction}]
%$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\begin{bmatrix} -\textsc{Cons} \\ -\textsc{Rnd} \\ +\textsc{Back} \\ -\textsc{High} \end{bmatrix}~\textrm{C}_0~\begin{bmatrix} +\textsc{Cons} \\ +\textsc{Labial} \end{bmatrix}~\textrm{C}_0~\gap{}$
$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\textrm{ɑ}~\textrm{C}_0~\begin{bmatrix} +\textsc{Cons} \\ +\textsc{Labial} \end{bmatrix}~\textrm{C}_0~\gap{}$
\end{example}
